# split-apply-combine
***

the combination of the function *split()* along with the function *lapply()*, or eventually *sapply()*, is a common strategy in R for calculating some statistic on a variable stratified by different goups;

the data structure is split into subsets defined by another variable, a function is applied over the subsets and the results are finally collated and returned as a new R object

```{r}
# define some data
x <- c(rnorm(10), runif(10))

# define some groups (factors)
f <- gl(2, 10)

# split the data over the groups (factors)
# to note that a list is returned by split()
x.f<-split(x, f)

# apply the mean to each element of the list
 lapply(x.f, mean)
```

the same approach is also possible with a dataframe

```{r}
s <- split(airquality, airquality$Month)
# lapply the function colMeans
lapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)})
# or alternatively sapply for a more compact result
sapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)})
```

split-apply with interaction (combination) of variables

```{r}
# define some data
x <- rnorm(12)
# define factor 1
f1 <- gl(4, 3)
# define factor 2
f2 <- gl(3, 4)
# interaction of variables (factors)
f1.f2<-interaction(f1, f2)
# split by dropping empty levels
x.f1.f2<-split(x, f1.f2, drop = TRUE)
# lapply the function mean
lapply(x.f1.f2, mean)
# or sapply the function mean
sapply(x.f1.f2, mean)

```



# apply to an array
***

```{r}
# define an array (2*2)*5 (i.e. 5 matrices 2*2 ) filled with normal random numbers
x<- array(rnorm(2*2*5),c(2,2,5))
x
```

```{r}
# apply the mean over the third dimension of the array
# i.e preserve first and second dimensions
apply(x, c(1,2), mean)
```


