# apply

****

```{r, echo=FALSE}
str(apply)
```

it does the following:

- evaluates a function (quite often an anonymous function) over the margins of an array or matrix (which is a special case of a 2*2 matrix) 

to note that second argument of *apply* (margin) indicates which dimension of the array (matrix) is going to be preserved (or retained)

```{r}
# define a matrix 10*5 filled with normal random numbers
x<- matrix(rnorm(50), 10, 5)
```

```{r}
# apply the mean on the rows (preserve rows, collapse columns)
apply(x, 1, mean)

# same result as above by using 
rowMeans(x) 

# apply the mean over the columns (preserve columns, collapse rows)
apply(x, 2, mean)

# same result as above by using 
colMeans(x) 

```

# lapply (list apply)
***

```{r, echo=FALSE}
str(lapply)
```

it does the following:

- loops over a list (iteration over each element in the list)
- applies a *function* to each element of a list
- returns a list

```{r}
# define a list
x<-list(a=1:10, b=rnorm(10), c=runif(10))
```

```{r}
# apply the function mean to all elements of the list
lapply(x, mean)
```

*lapply()* takes elements of the list and passes them as first argument of the function being applied: elements of sequence 1:3 are passed to the n argument of function *runif()*

```{r}
# evalute a function multiple times
lapply(1:3, runif)
```

*lapply()* accepts many arguments of the function passed as the main argument; i.e. any argument placed in the ... argument of function *lapply()* is passed down to the function which is applied to the single elements of the list

```{r}
# pass addtitional arguments of the function runif
lapply(1:3, runif, min=1, max=5)
```

*lapply()* very frequently makes use of anonymous functions: i.e. functions are
generated "on the fly" (once the call to *lapply()* is finished, the function
disappears and does not appear in the workspace)


# sapply (simplyfing list apply)

***

```{r, echo=FALSE}
str(sapply)
```

*sapply()* behaves similarly to *lapply()* with a difference in the returned value: sapply() tries to simplify the result whenever it is possible (if the result is a list where each element is of length 1 or a vector of the same length (>1), it returns a vector or a matrix;in other cases the result is again a list

```{r}
# compare with result of lapply()
sapply(x, mean)
```

# vapply (list apply returning a vector)

***

```{r, echo=FALSE}
str(vapply)
```

*vapply()* simplifies the result to be a vector or an array very much the same as *sapply()* but it takes a third argument specifying the template for  returned values

```{r}
# compare with result of lapply()
vapply(x, mean, numeric(1))
```

if the output does not fit the specified template, then *vapply()* will throw an error (which makes it less flexible than lapply, since the output must be the same size for each element and must be known in advance)

# tapply (apply to... tables?)

***

```{r, echo=FALSE}
str(tapply)
```

it is used to apply a function over subsets of a vector; it can be thought as a combination of *split()* and *sapply()* to vectors


```{r}
# define some data
x<- c(rnorm(10), runif(10), 1:10)
# define some groups (factors)
f<- gl(3,10)
```

```{r}
tapply(x, f, mean)
```

# mapply (multiple argument list apply)

***

```{r, echo=FALSE}
# it passes the sequence 1:3 to the first argument and the sequence 3:1 to the second argument of the function *rep()*
str(mapply)
```

it can be thought as multivariate *apply()* version, it applies a function in parallel over a set of defined arguments (it is used to iterate over multiple R object in parallel)

```{r}
mapply(rep, 1:3, 3:1)
```
*mapply()* behaves in the same way as sapply, by simplifying the output if it is possible; i can be used to "vectorize a function"

```{r}
# user defined function
sumsq<- function(m,s,x) {
  sum(((x-m)/s)^2)
}

# some data
x<- rnorm(100)

```


```{r}
mapply(sumsq,1:5,1:5, MoreArgs=list(x=x))
```

to note that very much the same result is obtained by calling the R function *Vectorize()* which creates a vectorized version of the user defined function

```{r}
vsumsq<- Vectorize(sumsq, c("m","s"))
vsumsq(1:5, 1:5, x)
```
